---
title: "Chapter 2: Sentencing"
output: html_notebook
---


Inputting Functions 
```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(lubridate)
library(scales)
library(tidyverse)
library(sf)
library(haven)
library(knitr)
install.packages("kableExtra")
library(data.table)
install.packages(c("readxl", "writexl"))
library("writexl")
library(broom)
install.packages("pander")
library(pander)
library(scales)
install.packages("estimatr")
library(estimatr)
library(kableExtra)
library(rvest)
library(stringr)
install.packages("webshot2")
library(webshot2)
install.packages("stargazer")
library(stargazer)
install.packages("WRS2")
install.packages("mc2d")
library(mc2d)
library(WRS2)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

Input Data for California
https://www.courts.ca.gov/38232.htm#felony-sentencing-trends

```{r}

California_County_Level_Sentences<- read.csv(
  "C:/Users/carol/Downloads/cjr-prop47-Dashboard-Data.csv",
  header = TRUE,
  check.names = TRUE
)

```


Definitions: 
Probation: allows individuals to remain in the community under county probation supervision for all or a portion of their sentence. They may serve up to one year in jail as a condition of their supervision. 
Straight and split sentences: include non-serious, non-violent, and non-sexual felony offenses. 
Straight sentence: an individual will serve their full term in county jail and then be released without supervision. 
Split sentence: an individual typically “splits” their term between jail time and community supervision. Occasionally individuals with a split sentence will serve their entire term supervised in the community with no jail time. 
Prison Sentence: the individual will be incarcerated, serving their sentence in state prison, and typically be released on parole or PRCS.
Since Realignment, individuals released from prison for non-serious, non-violent, and non-sexual offenses are typically supervised by county probation departments under Post Release Community Supervision (PRCS). Mandatory Supervision (MS) is the community supervision portion of a split sentence.
FTA Warrent = Presentence Warrants for Failure to Appear (FTAs) in Felony Cases


```{r}
California_County_Level_Sentences <- California_County_Level_Sentences %>%
  rename(
    Probation = Q4,
    Prison = Q3,
    Split_1170h = Q6,
    Straight_1170h = Q5,
    Probation_Petition = Q7,
    PRCS_Petition = Q18,
    MS_Petition = Q12,
    Parole_Petition = Q26,
    FTA_Warrent = Q2,
    Petition_for_Modification = Q1,
    FTA_Warrant = Q2,
    Petition_for_Modification = Q1)

California_County_Level_Sentences <- California_County_Level_Sentences %>%
  unite(County_and_Quarter, County, Quarter, remove=FALSE)

California_County_Level_Sentences




```
Upload CA DA data

```{r}

California_DA_election_data<- read.csv(
  "C:/Users/carol/Desktop/California_DAs_Excel.csv",
  header = TRUE,
  check.names = TRUE
)

```

Merge two dataframes and add percentages of sentences

```{r}



California_Merged_Data <- merge(California_County_Level_Sentences, California_DA_election_data, by="County_and_Quarter")

California_Merged_Data <- California_Merged_Data %>%
  mutate(Total_Sentences = Prison + Probation + Split_1170h + Straight_1170h)

California_Merged_Data <- California_Merged_Data %>%
  mutate(Percentage_Prison = Prison / Total_Sentences * 100,
           Percentage_Probation = Probation / Total_Sentences * 100 ,
         Percentage_Split = Split_1170h / Total_Sentences * 100,
         Percentage_Straight = Straight_1170h / Total_Sentences * 100)

write.csv(California_Merged_Data, "C:/Users/carol/Downloads/California_Merged_Data.csv", row.names=FALSE)


```


New York Counties 

```{r}

New_York_County_Level_Sentences <- read.csv(
  "C:/Users/carol/Desktop/NY_State_County_Level_Sentencing.csv",
  header = TRUE,
  check.names = TRUE
)
```

New York DA 

```{r}

New_York_DA <- read.csv(
  "C:/Users/carol/Desktop/NY_DA_Excel.csv",
  header = TRUE,
  check.names = TRUE
)

colnames(New_York_DA)

New_York_DA <- New_York_DA %>% 
       rename("County_and_Quarter" = "ï..County_and_Quarter",
              "Decarceratory" = "Decarceratory.",
              "Party" = "Party.",
            "Did_Incumbent_Seek_Reelection" =  "Did.Incumbent.Seek.Reelection.",
           "If_Sought_Reelection_Did_Incumbent_Win" = "If.Sought.Reelection..Did.Incumbent.Win." ,
           "Contested_Primary" = "Contested.Primary." ,
           "Contested_General" = "Contested.General." ,
           "Election_Year" = "Election.Year." )

```


Merge two dataframes

```{r}

NY_Merged_Data <- merge(New_York_County_Level_Sentences, New_York_DA, by="County_and_Quarter")

NY_Merged_Data <- NY_Merged_Data %>%
  mutate(Decarceratory = Decarceratory %>%
           as.factor,
         Election_Year = Election_Year %>%
           as.factor,
         Prison = Prison %>%
           as.numeric , 
         Probation = Probation %>%
           as.numeric , 
         Total = Total %>%
           as.numeric)

NY_Merged_Data <- NY_Merged_Data %>%
  mutate(Percentage_Prison = Prison / Total * 100,
           Percentage_Jail = Jail / Total * 100 ,
         Percentage_Jail_and_Probation = Jail.and.Probation / Total * 100,
         Percentage_Probation = Probation / Total * 100,
         Percentage_Fine = Fine / Total * 100,
         Percentage_Cond_Disc = Cond.Discharge / Total * 100)

write.csv(NY_Merged_Data, "C:/Users/carol/Downloads/NY_Merged_Data.csv", row.names=FALSE)

```



Pennsylvania 2013 Getting Proper Numbers

```{r}

Pennsylvania_2013_Sentences <- read.csv(
  "C:/Users/carol/Downloads/GEN_2013_NO_SID.csv",
  header = TRUE,
  check.names = TRUE
)


Pennsylvania_2013_Sentences <- Pennsylvania_2013_Sentences %>%
  mutate(quarter = DOS %>%
           as.POSIXct(format = "%m/%d/%Y") %>%
          quarter()) 

#next are all within the MS_SANCTION column

#prob(5), jail(3), prison(1)

Pennsylvania_Sanctions <- Pennsylvania_2013_Sentences %>%
  group_by(COUNTY, quarter, MS_SANCTION) %>%
  count()  

write.csv(Pennsylvania_Sanctions,  "C:/Users/carol/Downloads/Pennsylvania_Sanctions.csv", row.names=FALSE)

#conform(above, aggravated, below, mitigated, procedural, standard)

Pennsylvania_Conform <- Pennsylvania_2013_Sentences %>%
  group_by(COUNTY, quarter, CONFORM) %>%
  count() 



```

Pennsylvania 2014 Numbers

```{r}

Pennsylvania_2014_Sentences <- read.csv(
  "C:/Users/carol/Downloads/GEN_2014_NO_SID.csv",
  header = TRUE,
  check.names = TRUE
)


Pennsylvania_2014_Sentences <- Pennsylvania_2014_Sentences %>%
  mutate(quarter = DOS %>%
           as.POSIXct(format = "%m/%d/%Y") %>%
          quarter()) 

#next are all within the MS_SANCTION column

#prob(5), jail(3), prison(1)

Pennsylvania_Sanctions_2014 <- Pennsylvania_2014_Sentences %>%
  group_by(COUNTY, quarter, MS_SANCTION) %>%
  count()  

write.csv(Pennsylvania_Sanctions_2014,  "C:/Users/carol/Downloads/Pennsylvania_Sanctions_2014.csv", row.names=FALSE)

#conform(above, aggravated, below, mitigated, procedural, standard)

Pennsylvania_Conform_2014 <- Pennsylvania_2014_Sentences %>%
  group_by(COUNTY, quarter, CONFORM) %>%
  count() 


```

Pennsylvania 2015 Numbers

```{r}

Pennsylvania_2015_Sentences <- read.csv(
  "C:/Users/carol/Downloads/GEN_2015_NO_SID.csv",
  header = TRUE,
  check.names = TRUE
)


Pennsylvania_2015_Sentences <- Pennsylvania_2015_Sentences %>%
  mutate(quarter = DOS %>%
           as.POSIXct(format = "%m/%d/%Y") %>%
          quarter()) 

#next are all within the MS_SANCTION column

#prob(5), jail(3), prison(1)

Pennsylvania_Sanctions_2015 <- Pennsylvania_2015_Sentences %>%
  group_by(COUNTY, quarter, MS_SANCTION) %>%
  count()  

write.csv(Pennsylvania_Sanctions_2015,  "C:/Users/carol/Downloads/Pennsylvania_Sanctions_2015.csv", row.names=FALSE)

#conform(above, aggravated, below, mitigated, procedural, standard)

Pennsylvania_Conform_2015 <- Pennsylvania_2015_Sentences %>%
  group_by(COUNTY, quarter, CONFORM) %>%
  count() 


```

Pennsylvania 2016 Numbers

```{r}

Pennsylvania_2016_Sentences <- read.csv(
  "C:/Users/carol/Downloads/GEN_2016_NO_SID.csv",
  header = TRUE,
  check.names = TRUE
)


Pennsylvania_2016_Sentences <- Pennsylvania_2016_Sentences %>%
  mutate(quarter = DOS %>%
           as.POSIXct(format = "%m/%d/%Y") %>%
          quarter()) 

#next are all within the MS_SANCTION column

#prob(5), jail(3), prison(1)

Pennsylvania_Sanctions_2016 <- Pennsylvania_2016_Sentences %>%
  group_by(COUNTY, quarter, MS_SANCTION) %>%
  count()  

write.csv(Pennsylvania_Sanctions_2016,  "C:/Users/carol/Downloads/Pennsylvania_Sanctions_2016.csv", row.names=FALSE)

#conform(above, aggravated, below, mitigated, procedural, standard)

Pennsylvania_Conform_2016 <- Pennsylvania_2016_Sentences %>%
  group_by(COUNTY, quarter, CONFORM) %>%
  count() 


```


```{r}

Pennsylvania_2017_Sentences <- read.csv(
  "C:/Users/carol/Downloads/GEN_2017_NO_SID.csv",
  header = TRUE,
  check.names = TRUE
)


Pennsylvania_2017_Sentences <- Pennsylvania_2017_Sentences %>%
  mutate(quarter = DOS %>%
           as.POSIXct(format = "%m/%d/%Y") %>%
          quarter()) 

#next are all within the MS_SANCTION column

#prob(5), jail(3), prison(1)

Pennsylvania_Sanctions_2017 <- Pennsylvania_2017_Sentences %>%
  group_by(COUNTY, quarter, MS_SANCTION) %>%
  count()  

write.csv(Pennsylvania_Sanctions_2017,  "C:/Users/carol/Downloads/Pennsylvania_Sanctions_2017.csv", row.names=FALSE)

#conform(above, aggravated, below, mitigated, procedural, standard)

Pennsylvania_Conform_2017 <- Pennsylvania_2017_Sentences %>%
  group_by(COUNTY, quarter, CONFORM) %>%
  count() 

Pennsylvania_2018_Sentences <- read.csv(
  "C:/Users/carol/Downloads/GEN_2018_NO_SID.csv",
  header = TRUE,
  check.names = TRUE
)


Pennsylvania_2018_Sentences <- Pennsylvania_2018_Sentences %>%
  mutate(quarter = DOS %>%
           as.POSIXct(format = "%m/%d/%Y") %>%
          quarter()) 

#next are all within the MS_SANCTION column

#prob(5), jail(3), prison(1)

Pennsylvania_Sanctions_2018 <- Pennsylvania_2018_Sentences %>%
  group_by(COUNTY, quarter, MS_SANCTION) %>%
  count()  

write.csv(Pennsylvania_Sanctions_2018,  "C:/Users/carol/Downloads/Pennsylvania_Sanctions_2018.csv", row.names=FALSE)

#conform(above, aggravated, below, mitigated, procedural, standard)

Pennsylvania_Conform_2018 <- Pennsylvania_2018_Sentences %>%
  group_by(COUNTY, quarter, CONFORM) %>%
  count() 

Pennsylvania_2019_Sentences <- read.csv(
  "C:/Users/carol/Downloads/GEN_2019_NO_SID.csv",
  header = TRUE,
  check.names = TRUE
)

str(Pennsylvania_2019_Sentences$DOS)

#this is not working. Quarter is NA here.


Pennsylvania_2019_Sentences <- Pennsylvania_2019_Sentences %>%
  mutate(quarter = DOS %>%
           as.POSIXct(format = "%d-%b-%y") %>%
           quarter())


#next are all within the MS_SANCTION column

#prob(5), jail(3), prison(1)

Pennsylvania_Sanctions_2019 <- Pennsylvania_2019_Sentences %>%
  group_by(COUNTY, quarter, MS_SANCTION) %>%
  count()  

write.csv(Pennsylvania_Sanctions_2019,  "C:/Users/carol/Downloads/Pennsylvania_Sanctions_2019.csv", row.names=FALSE)

#conform(above, aggravated, below, mitigated, procedural, standard)

Pennsylvania_Conform_2019 <- Pennsylvania_2019_Sentences %>%
  group_by(COUNTY, quarter, CONFORM) %>%
  count() 

Pennsylvania_2020_Sentences <- read.csv(
  "C:/Users/carol/Downloads/GEN_2020_NO_SID.csv",
  header = TRUE,
  check.names = TRUE
)


Pennsylvania_2020_Sentences <- Pennsylvania_2020_Sentences %>%
  mutate(quarter = DOS %>%
           as.POSIXct(format = "%d-%b-%y") %>%
          quarter()) 

#next are all within the MS_SANCTION column

#prob(5), jail(3), prison(1)

Pennsylvania_Sanctions_2020 <- Pennsylvania_2020_Sentences %>%
  group_by(COUNTY, quarter, MS_SANCTION) %>%
  count()  

write.csv(Pennsylvania_Sanctions_2020,  "C:/Users/carol/Downloads/Pennsylvania_Sanctions_2020.csv", row.names=FALSE)

Pennsylvania_2021_Sentences <- read.csv(
  "C:/Users/carol/Downloads/GEN_2021_NO_SID.csv",
  header = TRUE,
  check.names = TRUE
)


Pennsylvania_2021_Sentences <- Pennsylvania_2021_Sentences %>%
  mutate(quarter = DOS %>%
           as.POSIXct(format = "%d-%b-%y") %>%
          quarter()) 

#next are all within the MS_SANCTION column

#prob(5), jail(3), prison(1)

Pennsylvania_Sanctions_2021 <- Pennsylvania_2021_Sentences %>%
  group_by(COUNTY, quarter, MS_SANCTION) %>%
  count()  

write.csv(Pennsylvania_Sanctions_2021,  "C:/Users/carol/Downloads/Pennsylvania_Sanctions_2021.csv", row.names=FALSE)



```


```{r}
PA_State_County_Level_Sentencing <- read.csv(
  "C:/Users/carol/Desktop/PA_State_County_Level_Sentencing.csv",
  header = TRUE,
  check.names = TRUE
)

PA_State_County_Level_Sentencing_with_Year <- PA_State_County_Level_Sentencing %>%
  mutate(Year = case_when(str_detect(County_and_Quarter, "2013") ~ "2013", 
                         str_detect(County_and_Quarter, "2014") ~ "2014",
                         str_detect(County_and_Quarter, "2015") ~ "2015",
                         str_detect(County_and_Quarter, "2016") ~ "2016",
                         str_detect(County_and_Quarter, "2017") ~ "2017",
                         str_detect(County_and_Quarter, "2018") ~ "2018",
                         str_detect(County_and_Quarter, "2019") ~ "2019",
                         str_detect(County_and_Quarter, "2020") ~ "2020",
                         str_detect(County_and_Quarter, "2021") ~ "2021"))


write.csv(PA_State_County_Level_Sentencing_with_Year,  "C:/Users/carol/Downloads/PA_State_County_Level_Sentencing_with_Year.csv", row.names = FALSE)


```

Merging PA Info

```{r}
PA_Sentencing <- read.csv(
  "C:/Users/carol/Downloads/PA_Sentencing.csv",
  header = TRUE,
  check.names = TRUE
)

#prob(5), jail(3), prison(1)

PA_Sentencing <- PA_Sentencing %>%
  rename("County_Jail" = "X3",
         "State_Prison" = "X1",
         "Probation" = "X5",
         "Restorative_Sanction" = "X6",
         "Restrictive_intermediate_punishment" = "X4", 
         "State_intermediate_punishment" = "X2")

PA_DAs <- read.csv(
  "C:/Users/carol/Downloads/PA_DA.csv",
  header = TRUE,
  check.names = TRUE
)

PA_Merged_Data <- merge(PA_Sentencing, PA_DAs, by="County_and_Quarter")

write.csv(PA_Merged_Data, "C:/Users/carol/Downloads/PA_Merged_Data.csv", row.names = FALSE)

```

Load all cleaned data

```{r}

NY_Merged_Data<- read.csv(
  "C:/Users/carol/Downloads/NY_Merged_Data.csv",
  header = TRUE,
  check.names = TRUE
)

PA_Merged_Data<- read.csv(
  "C:/Users/carol/Downloads/PA_Merged_Data.csv",
  header = TRUE,
  check.names = TRUE
)

CA_Merged_Data <- read.csv(
  "C:/Users/carol/Downloads/CA_Merged_Data.csv",
  header = TRUE,
  check.names = TRUE
)

```

Analyzing NY Data

```{r}


NY_Prison_Sentences_by_Decarceratory <- aov(data = NY_Merged_Data , Percentage_Prison ~ Decarceratory)

summary(NY_Prison_Sentences_by_Decarceratory)


TukeyHSD(NY_Prison_Sentences_by_Decarceratory)


wilcox.test(NY_Merged_Data %>%
              filter(Decarceratory == 1) %>%
              select(Percentage_Prison) %>%
              na.omit %>%
              unlist, 
            NY_Merged_Data %>%
              filter(Decarceratory == 0) %>%
              select(Percentage_Prison) %>%
              na.omit %>%
              unlist)

#still comparing medians

med1way(data = NY_Merged_Data ,
        Percentage_Prison ~ Decarceratory)

med2way(data = NY_Merged_Data ,
        Percentage_Prison ~ Decarceratory + Election_Year)

#boxplot

NY_Merged_Data %>%
  ggplot(aes(fill = Election_Year, x = Decarceratory , y = Percentage_Prison)) + geom_boxplot() + facet_wrap(~ Election_Year) 

#is there a correlation between probation and prison

cor.test(data = NY_Merged_Data , 
         ~ Prison + Probation)

#graph above

NY_Merged_Data %>%
  ggplot(aes(color = Party , x = Prison , y = Probation)) + geom_point()

```

Analyzing CA Data
```{r}

CA_Prison_Sentences_by_Decarceratory <- aov(data = CA_Merged_Data , Percentage_Prison ~ Decarceratory.)

summary(CA_Prison_Sentences_by_Decarceratory)


TukeyHSD(CA_Prison_Sentences_by_Decarceratory)


wilcox.test(CA_Merged_Data %>%
              filter(Decarceratory. == 1) %>%
              select(Percentage_Prison) %>%
              na.omit %>%
              unlist, 
            CA_Merged_Data %>%
              filter(Decarceratory. == 0) %>%
              select(Percentage_Prison) %>%
              na.omit %>%
              unlist)

#still comparing medians

med1way(data = CA_Merged_Data ,
        Percentage_Prison ~ Decarceratory.)

med2way(data = CA_Merged_Data ,
        Percentage_Prison ~ Decarceratory. + Election.Year.)

#boxplot

CA_Merged_Data %>%
  ggplot(aes(fill = Election.Year., x = Decarceratory. , y = Percentage_Prison)) + geom_boxplot() + facet_wrap(~ Election.Year.) 

#is there a correlation between probation and prison

cor.test(data = v , 
         ~ Percentage_Prison + Percentage_Probation)

#graph above

CA_Merged_Data %>%
  ggplot(aes(color = Decarceratory. , x = Prison , y = Probation)) + geom_point()

```
```

