---
title: "Chapter 2: Sentencing"
output: html_notebook
---


Inputting Functions 
```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(lubridate)
library(scales)
library(tidyverse)
library(sf)
library(haven)
library(knitr)
library(data.table)
library("writexl")
library(broom)
library(pander)
library(scales)
library(estimatr)
library(kableExtra)
library(rvest)
library(stringr)
library(webshot2)
library(stargazer)
library(mc2d)
library(WRS2)
library(ggsignif)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

Input Data for California
https://www.courts.ca.gov/38232.htm#felony-sentencing-trends

```{r}

California_County_Level_Sentences<- read.csv(
  "C:/Users/carol/Downloads/cjr-prop47-Dashboard-Data.csv",
  header = TRUE,
  check.names = TRUE
)

```


Definitions: 
Probation: allows individuals to remain in the community under county probation supervision for all or a portion of their sentence. They may serve up to one year in jail as a condition of their supervision. 
Straight and split sentences: include non-serious, non-violent, and non-sexual felony offenses. 
Straight sentence: an individual will serve their full term in county jail and then be released without supervision. 
Split sentence: an individual typically “splits” their term between jail time and community supervision. Occasionally individuals with a split sentence will serve their entire term supervised in the community with no jail time. 
Prison Sentence: the individual will be incarcerated, serving their sentence in state prison, and typically be released on parole or PRCS.
Since Realignment, individuals released from prison for non-serious, non-violent, and non-sexual offenses are typically supervised by county probation departments under Post Release Community Supervision (PRCS). Mandatory Supervision (MS) is the community supervision portion of a split sentence.
FTA Warrent = Presentence Warrants for Failure to Appear (FTAs) in Felony Cases


```{r}
California_County_Level_Sentences <- California_County_Level_Sentences %>%
  rename(
    Probation = Q4,
    Prison = Q3,
    Split_1170h = Q6,
    Straight_1170h = Q5,
    Probation_Petition = Q7,
    PRCS_Petition = Q18,
    MS_Petition = Q12,
    Parole_Petition = Q26,
    FTA_Warrent = Q2,
    Petition_for_Modification = Q1,
    FTA_Warrant = Q2,
    Petition_for_Modification = Q1)

California_County_Level_Sentences <- California_County_Level_Sentences %>%
  unite(County_and_Quarter, County, Quarter, remove=FALSE)

California_County_Level_Sentences




```
Upload CA DA data

```{r}

California_DA_election_data<- read.csv(
  "C:/Users/carol/Desktop/California_DAs_Excel.csv",
  header = TRUE,
  check.names = TRUE
)


#Merge two dataframes and add percentages of sentences

California_Merged_Data <- merge(California_County_Level_Sentences, California_DA_election_data, by="County_and_Quarter")

California_Merged_Data <- California_Merged_Data %>%
  mutate(Total_Sentences = Prison + Probation + Split_1170h + Straight_1170h)

California_Merged_Data <- California_Merged_Data %>%
  mutate(Percentage_Prison = Prison / Total_Sentences * 100,
           Percentage_Probation = Probation / Total_Sentences * 100 ,
         Percentage_Split = Split_1170h / Total_Sentences * 100,
         Percentage_Straight = Straight_1170h / Total_Sentences * 100)

write.csv(CA_Merged_Data, "C:/Users/carol/Downloads/California_Merged_Data.csv", row.names=FALSE)


```


New York Counties 

```{r}

New_York_County_Level_Sentences <- read.csv(
  "C:/Users/carol/Desktop/NY_State_County_Level_Sentencing.csv",
  header = TRUE,
  check.names = TRUE
)
```

New York DA 

```{r}

New_York_DA <- read.csv(
  "C:/Users/carol/Desktop/NY_DA_Excel.csv",
  header = TRUE,
  check.names = TRUE
)

colnames(New_York_DA)

New_York_DA <- New_York_DA %>% 
       rename("County_and_Quarter" = "ï..County_and_Quarter",
              "Decarceratory" = "Decarceratory.",
              "Party" = "Party.",
            "Did_Incumbent_Seek_Reelection" =  "Did.Incumbent.Seek.Reelection.",
           "If_Sought_Reelection_Did_Incumbent_Win" = "If.Sought.Reelection..Did.Incumbent.Win." ,
           "Contested_Primary" = "Contested.Primary." ,
           "Contested_General" = "Contested.General." ,
           "Election_Year" = "Election.Year." )

```


Merge two dataframes

```{r}

NY_Merged_Data <- merge(New_York_County_Level_Sentences, New_York_DA, by="County_and_Quarter")

NY_Merged_Data <- NY_Merged_Data %>%
  mutate(Decarceratory = Decarceratory %>%
           as.factor,
         Election_Year = Election_Year %>%
           as.factor,
         Prison = Prison %>%
           as.numeric , 
         Probation = Probation %>%
           as.numeric , 
         Total = Total %>%
           as.numeric)

NY_Merged_Data <- NY_Merged_Data %>%
  mutate(Percentage_Prison = Prison / Total * 100,
           Percentage_Jail = Jail / Total * 100 ,
         Percentage_Jail_and_Probation = Jail.and.Probation / Total * 100,
         Percentage_Probation = Probation / Total * 100,
         Percentage_Fine = Fine / Total * 100,
         Percentage_Cond_Disc = Cond.Discharge / Total * 100)



write.csv(NY_Merged_Data, "C:/Users/carol/Downloads/NY_Merged_Data.csv", row.names=FALSE)

```



Pennsylvania 2013 Getting Proper Numbers

```{r}

Pennsylvania_2013_Sentences <- read.csv(
  "C:/Users/carol/Downloads/GEN_2013_NO_SID.csv",
  header = TRUE,
  check.names = TRUE
)


Pennsylvania_2013_Sentences <- Pennsylvania_2013_Sentences %>%
  mutate(quarter = DOS %>%
           as.POSIXct(format = "%m/%d/%Y") %>%
          quarter()) 

#next are all within the MS_SANCTION column

#prob(5), jail(3), prison(1)

Pennsylvania_Sanctions <- Pennsylvania_2013_Sentences %>%
  group_by(COUNTY, quarter, MS_SANCTION) %>%
  count()  

write.csv(Pennsylvania_Sanctions,  "C:/Users/carol/Downloads/Pennsylvania_Sanctions.csv", row.names=FALSE)

#conform(above, aggravated, below, mitigated, procedural, standard)

Pennsylvania_Conform <- Pennsylvania_2013_Sentences %>%
  group_by(COUNTY, quarter, CONFORM) %>%
  count() 



```

Pennsylvania 2014 Numbers

```{r}

Pennsylvania_2014_Sentences <- read.csv(
  "C:/Users/carol/Downloads/GEN_2014_NO_SID.csv",
  header = TRUE,
  check.names = TRUE
)


Pennsylvania_2014_Sentences <- Pennsylvania_2014_Sentences %>%
  mutate(quarter = DOS %>%
           as.POSIXct(format = "%m/%d/%Y") %>%
          quarter()) 

#next are all within the MS_SANCTION column

#prob(5), jail(3), prison(1)

Pennsylvania_Sanctions_2014 <- Pennsylvania_2014_Sentences %>%
  group_by(COUNTY, quarter, MS_SANCTION) %>%
  count()  

write.csv(Pennsylvania_Sanctions_2014,  "C:/Users/carol/Downloads/Pennsylvania_Sanctions_2014.csv", row.names=FALSE)

#conform(above, aggravated, below, mitigated, procedural, standard)

Pennsylvania_Conform_2014 <- Pennsylvania_2014_Sentences %>%
  group_by(COUNTY, quarter, CONFORM) %>%
  count() 


```

Pennsylvania 2015 Numbers

```{r}

Pennsylvania_2015_Sentences <- read.csv(
  "C:/Users/carol/Downloads/GEN_2015_NO_SID.csv",
  header = TRUE,
  check.names = TRUE
)


Pennsylvania_2015_Sentences <- Pennsylvania_2015_Sentences %>%
  mutate(quarter = DOS %>%
           as.POSIXct(format = "%m/%d/%Y") %>%
          quarter()) 

#next are all within the MS_SANCTION column

#prob(5), jail(3), prison(1)

Pennsylvania_Sanctions_2015 <- Pennsylvania_2015_Sentences %>%
  group_by(COUNTY, quarter, MS_SANCTION) %>%
  count()  

write.csv(Pennsylvania_Sanctions_2015,  "C:/Users/carol/Downloads/Pennsylvania_Sanctions_2015.csv", row.names=FALSE)

#conform(above, aggravated, below, mitigated, procedural, standard)

Pennsylvania_Conform_2015 <- Pennsylvania_2015_Sentences %>%
  group_by(COUNTY, quarter, CONFORM) %>%
  count() 


```

Pennsylvania 2016 Numbers

```{r}

Pennsylvania_2016_Sentences <- read.csv(
  "C:/Users/carol/Downloads/GEN_2016_NO_SID.csv",
  header = TRUE,
  check.names = TRUE
)


Pennsylvania_2016_Sentences <- Pennsylvania_2016_Sentences %>%
  mutate(quarter = DOS %>%
           as.POSIXct(format = "%m/%d/%Y") %>%
          quarter()) 

#next are all within the MS_SANCTION column

#prob(5), jail(3), prison(1)

Pennsylvania_Sanctions_2016 <- Pennsylvania_2016_Sentences %>%
  group_by(COUNTY, quarter, MS_SANCTION) %>%
  count()  

write.csv(Pennsylvania_Sanctions_2016,  "C:/Users/carol/Downloads/Pennsylvania_Sanctions_2016.csv", row.names=FALSE)

#conform(above, aggravated, below, mitigated, procedural, standard)

Pennsylvania_Conform_2016 <- Pennsylvania_2016_Sentences %>%
  group_by(COUNTY, quarter, CONFORM) %>%
  count() 


```


```{r}

Pennsylvania_2017_Sentences <- read.csv(
  "C:/Users/carol/Downloads/GEN_2017_NO_SID.csv",
  header = TRUE,
  check.names = TRUE
)


Pennsylvania_2017_Sentences <- Pennsylvania_2017_Sentences %>%
  mutate(quarter = DOS %>%
           as.POSIXct(format = "%m/%d/%Y") %>%
          quarter()) 

#next are all within the MS_SANCTION column

#prob(5), jail(3), prison(1)

Pennsylvania_Sanctions_2017 <- Pennsylvania_2017_Sentences %>%
  group_by(COUNTY, quarter, MS_SANCTION) %>%
  count()  

write.csv(Pennsylvania_Sanctions_2017,  "C:/Users/carol/Downloads/Pennsylvania_Sanctions_2017.csv", row.names=FALSE)

#conform(above, aggravated, below, mitigated, procedural, standard)

Pennsylvania_Conform_2017 <- Pennsylvania_2017_Sentences %>%
  group_by(COUNTY, quarter, CONFORM) %>%
  count() 

Pennsylvania_2018_Sentences <- read.csv(
  "C:/Users/carol/Downloads/GEN_2018_NO_SID.csv",
  header = TRUE,
  check.names = TRUE
)


Pennsylvania_2018_Sentences <- Pennsylvania_2018_Sentences %>%
  mutate(quarter = DOS %>%
           as.POSIXct(format = "%m/%d/%Y") %>%
          quarter()) 

#next are all within the MS_SANCTION column

#prob(5), jail(3), prison(1)

Pennsylvania_Sanctions_2018 <- Pennsylvania_2018_Sentences %>%
  group_by(COUNTY, quarter, MS_SANCTION) %>%
  count()  

write.csv(Pennsylvania_Sanctions_2018,  "C:/Users/carol/Downloads/Pennsylvania_Sanctions_2018.csv", row.names=FALSE)

#conform(above, aggravated, below, mitigated, procedural, standard)

Pennsylvania_Conform_2018 <- Pennsylvania_2018_Sentences %>%
  group_by(COUNTY, quarter, CONFORM) %>%
  count() 

Pennsylvania_2019_Sentences <- read.csv(
  "C:/Users/carol/Downloads/GEN_2019_NO_SID.csv",
  header = TRUE,
  check.names = TRUE
)

str(Pennsylvania_2019_Sentences$DOS)

#this is not working. Quarter is NA here.


Pennsylvania_2019_Sentences <- Pennsylvania_2019_Sentences %>%
  mutate(quarter = DOS %>%
           as.POSIXct(format = "%d-%b-%y") %>%
           quarter())


#next are all within the MS_SANCTION column

#prob(5), jail(3), prison(1)

Pennsylvania_Sanctions_2019 <- Pennsylvania_2019_Sentences %>%
  group_by(COUNTY, quarter, MS_SANCTION) %>%
  count()  

write.csv(Pennsylvania_Sanctions_2019,  "C:/Users/carol/Downloads/Pennsylvania_Sanctions_2019.csv", row.names=FALSE)

#conform(above, aggravated, below, mitigated, procedural, standard)

Pennsylvania_Conform_2019 <- Pennsylvania_2019_Sentences %>%
  group_by(COUNTY, quarter, CONFORM) %>%
  count() 

Pennsylvania_2020_Sentences <- read.csv(
  "C:/Users/carol/Downloads/GEN_2020_NO_SID.csv",
  header = TRUE,
  check.names = TRUE
)


Pennsylvania_2020_Sentences <- Pennsylvania_2020_Sentences %>%
  mutate(quarter = DOS %>%
           as.POSIXct(format = "%d-%b-%y") %>%
          quarter()) 

#next are all within the MS_SANCTION column

#prob(5), jail(3), prison(1)

Pennsylvania_Sanctions_2020 <- Pennsylvania_2020_Sentences %>%
  group_by(COUNTY, quarter, MS_SANCTION) %>%
  count()  

write.csv(Pennsylvania_Sanctions_2020,  "C:/Users/carol/Downloads/Pennsylvania_Sanctions_2020.csv", row.names=FALSE)

Pennsylvania_2021_Sentences <- read.csv(
  "C:/Users/carol/Downloads/GEN_2021_NO_SID.csv",
  header = TRUE,
  check.names = TRUE
)


Pennsylvania_2021_Sentences <- Pennsylvania_2021_Sentences %>%
  mutate(quarter = DOS %>%
           as.POSIXct(format = "%d-%b-%y") %>%
          quarter()) 

#next are all within the MS_SANCTION column

#prob(5), jail(3), prison(1)

Pennsylvania_Sanctions_2021 <- Pennsylvania_2021_Sentences %>%
  group_by(COUNTY, quarter, MS_SANCTION) %>%
  count()  

write.csv(Pennsylvania_Sanctions_2021,  "C:/Users/carol/Downloads/Pennsylvania_Sanctions_2021.csv", row.names=FALSE)



```


```{r}
PA_State_County_Level_Sentencing <- read.csv(
  "C:/Users/carol/Desktop/PA_State_County_Level_Sentencing.csv",
  header = TRUE,
  check.names = TRUE
)

PA_State_County_Level_Sentencing_with_Year <- PA_State_County_Level_Sentencing %>%
  mutate(Year = case_when(str_detect(County_and_Quarter, "2013") ~ "2013", 
                         str_detect(County_and_Quarter, "2014") ~ "2014",
                         str_detect(County_and_Quarter, "2015") ~ "2015",
                         str_detect(County_and_Quarter, "2016") ~ "2016",
                         str_detect(County_and_Quarter, "2017") ~ "2017",
                         str_detect(County_and_Quarter, "2018") ~ "2018",
                         str_detect(County_and_Quarter, "2019") ~ "2019",
                         str_detect(County_and_Quarter, "2020") ~ "2020",
                         str_detect(County_and_Quarter, "2021") ~ "2021"))


write.csv(PA_State_County_Level_Sentencing_with_Year,  "C:/Users/carol/Downloads/PA_State_County_Level_Sentencing_with_Year.csv", row.names = FALSE)


```

Merging PA Info

```{r}
PA_Sentencing <- read.csv(
  "C:/Users/carol/Downloads/PA_Sentencing.csv",
  header = TRUE,
  check.names = TRUE
)

#prob(5), jail(3), prison(1)

PA_Sentencing <- PA_Sentencing %>%
  rename("County_Jail" = "X3",
         "State_Prison" = "X1",
         "Probation" = "X5",
         "Restorative_Sanction" = "X6",
         "Restrictive_intermediate_punishment" = "X4", 
         "State_intermediate_punishment" = "X2")

PA_DAs <- read.csv(
  "C:/Users/carol/Downloads/PA_DA.csv",
  header = TRUE,
  check.names = TRUE
)

PA_Merged_Data <- merge(PA_Sentencing, PA_DAs, by="County_and_Quarter")

PA_Merged_Data <- PA_Merged_Data %>%
  mutate(Decarceratory = Decarceratory. %>%
           as.factor,
         Election_Year = Election.Year. %>%
           as.factor,
         State_Prison = State_Prison %>%
           as.numeric , 
         Probation = Probation %>%
           as.numeric , 
         Total = Total %>%
           as.numeric ,
         County_Jail = County_Jail %>%
           as.numeric ,
         Restrictive_intermediate_punishment = Restrictive_intermediate_punishment %>%
           as.numeric ,
         Restorative_Sanction = Restorative_Sanction %>%
           as.numeric , 
         State_intermediate_punishment = State_intermediate_punishment %>%
           as.numeric)

PA_Merged_Data <- PA_Merged_Data %>%
  mutate(Percentage_Prison = State_Prison / Total * 100,
           Percentage_Jail = County_Jail / Total * 100 ,
         Percentage_Intermediate_Punishment = State_intermediate_punishment / Total * 100,
         Percentage_Probation = Probation / Total * 100,
         Percentage_Restorative_Sanction = Restorative_Sanction / Total * 100)

colnames(PA_Merged_Data)[2]<-"County"
colnames(PA_Merged_Data)[11]<-"County" 

PA_Merged_Data <- PA_Merged_Data %>%
  mutate(Election.Year. = ifelse(is.na(Election.Year.), 0, Election.Year.))

PA_Merged_Data <- PA_Merged_Data %>% select(-Election.Year.)

PA_Merged_Data <- PA_Merged_Data %>%
  mutate(
         Election_Year = Election.Year. %>%
           as.factor)

write.csv(PA_Merged_Data, "C:/Users/carol/Downloads/PA_Merged_Data.csv", row.names=FALSE)

```

Load all cleaned data

```{r}

NY_Merged_Data<- read.csv(
  "C:/Users/carol/Downloads/NY_Merged_Data.csv",
  header = TRUE,
  check.names = F
)

PA_Merged_Data<- read.csv(
  "C:/Users/carol/Downloads/PA_Merged_Data.csv",
  header = TRUE,
  check.names = F
)

CA_Merged_Data <- read.csv(
  "C:/Users/carol/Downloads/CA_Merged_Data.csv",
  header = TRUE,
  check.names = F
)

```

Analyzing NY Data

```{r}
NY_Merged_Data <- NY_Merged_Data %>%
  mutate(Decarceratory = Decarceratory %>%
           as.factor,
         Election_Year = Election_Year %>%
           as.factor) 

NY_Prison_Sentences_by_Decarceratory <- lm(data = NY_Merged_Data , Percentage_Prison ~ Decarceratory)

class(NY_Prison_Sentences_by_Decarceratory)


stargazer(NY_Prison_Sentences_by_Decarceratory,  title = "P-Test Results") %>%
  write_lines("NY_Prison_Sent_By_Dec.txt")

NY_Prison_Sentences_by_Decarceratory_aov <- aov(data = NY_Merged_Data , Percentage_Prison ~ Decarceratory)


TurkeyHSD_NY_Prison_Sentences <- TukeyHSD(NY_Prison_Sentences_by_Decarceratory_aov)

TurkeyHSD_NY_Prison_Sentences

tidy_turkey(TurkeyHSD_NY_Prison_Sentences)

#look at the means to see the difference


summary(NY_Prison_Sentences_by_Decarceratory)

#ggplot it

ggplot(data = NY_Merged_Data , aes(Decarceratory, Percentage_Prison)) + 
  geom_boxplot()

NY_Merged_Data %>%
drop_na(Decarceratory) %>%
  ggplot(aes(Decarceratory, Percentage_Prison)) +
  geom_boxplot()

#wilcox test compares medians 

wilcox_test_decarceratory <- wilcox.test(NY_Merged_Data %>%
              filter(Decarceratory == 1) %>%
              select(Percentage_Prison) %>%
              na.omit %>%
              unlist, 
            NY_Merged_Data %>%
              filter(Decarceratory == 0) %>%
              select(Percentage_Prison) %>%
              na.omit %>%
              unlist)


#still comparing medians. p values on the two main effects and the chi-square is the interaction between the two. Chi-square is a test of independence. So if two things are when you reject the null hypothesis you are saying that you cannot reject the null hypothesis that they are independent. So we do not have evidence showing dependence. 

med1way(data = NY_Merged_Data ,
        Percentage_Prison ~ Decarceratory)

med2way_Dec_Election_Year <- med2way(data = NY_Merged_Data ,
        Percentage_Prison ~ Decarceratory + Election_Year)

med2way_Dec_Election_Year
summary(med2way_Dec_Election_Year)

#mean comparison

AOV_Mean_Dec_Election <- aov(data = NY_Merged_Data , Percentage_Prison ~ Decarceratory * Election_Year)

summary(AOV_Mean_Dec_Election)

anova(AOV_Mean_Dec_Election)
#boxplot 
#http://sthda.com/english/articles/24-ggpubr-publication-ready-plots/76-add-p-values-and-significance-levels-to-ggplots/

#https://www.datanovia.com/en/lessons/kruskal-wallis-test-in-r/

#work on fixing this plot to be easier to read 

library(palmerpenguins)
data(package = 'palmerpenguins')


NY_Merged_Data %>%
  drop_na(Decarceratory) %>%
  ggplot(aes(x = Decarceratory , y = Percentage_Prison)) + 
  geom_boxplot(aes(fill = Election_Year)) + 
  facet_wrap(~ Election_Year) +
  geom_signif(comparisons = list(c("0" , "1")) , 
    map_signif_level = TRUE,
    textsize = 6,
    step_increase = .1
  ) +
  ylim(NA , 80) +
  labs(
    x = "Decarceratory",
    y = "Percentage of Guilty Verdicts Sentenced to Prison",
    title = "Differences in Prison Sentencing Percentage by Election Year",
    caption = "Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1"
  ) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))


s#is there a correlation between probation and prison

cor.test(data = NY_Merged_Data , 
         ~ Percentage_Prison + Percentage_Probation)

#graph above

NY_Merged_Data %>%
  ggplot(aes(color = Party , x = Percentage_Prison , y = Percentage_Probation)) + geom_point()

# create an independent and dependent variable. Think about if you want to do a regression (y-intercept) or Anova

NY_Merged_Data_Model_Jail <-  
  lm(Percentage_Jail ~ Contested_Primary, data = NY_Merged_Data) 


NY_Merged_Data_Model_Probation <- 
  lm(Percentage_Probation ~ Contested_Primary, data = NY_Merged_Data)


NY_Merged_Data_Model_Fine <- 
  lm(Percentage_Fine ~ Contested_Primary, data = NY_Merged_Data)


NY_Merged_Data_Model_Prison <- 
  lm(Percentage_Prison ~ Contested_Primary, data = NY_Merged_Data)

#slope = estimate of contested_primary

# y = 6.95  - 1.29x 

summary(NY_Merged_Data_Model_Prison)
summary(NY_Merged_Data_Model_Fine)
summary(NY_Merged_Data_Model_Probation)
summary(NY_Merged_Data_Model_Jail)



```

Analyzing CA Data
```{r}

CA_Merged_Data <- CA_Merged_Data %>%
  mutate(Decarceratory. = Decarceratory. %>%
           as.factor,
         Election.Year. = Election.Year. %>%
           as.factor) 

CA_Prison_Sentences_by_Decarceratory <- aov(data = CA_Merged_Data , Percentage_Prison ~ Decarceratory.)

summary(CA_Prison_Sentences_by_Decarceratory)


TukeyHSD(CA_Prison_Sentences_by_Decarceratory)

#the above might be interesting because it's not a partisan election? I am not totally sure if that actually tells us anything

wilcox.test(CA_Merged_Data %>%
              filter(Decarceratory. == 1) %>%
              select(Percentage_Prison) %>%
              na.omit %>%
              unlist, 
            CA_Merged_Data %>%
              filter(Decarceratory. == 0) %>%
              select(Percentage_Prison) %>%
              na.omit %>%
              unlist)

#still comparing medians

med1way(data = CA_Merged_Data ,
        Percentage_Prison ~ Decarceratory.)

med2way(data = CA_Merged_Data ,
        Percentage_Prison ~ Decarceratory. + Election.Year.)

#boxplot

CA_Merged_Data %>%
  ggplot(aes(fill = Election.Year., x = Decarceratory. , y = Percentage_Prison)) + geom_boxplot() + facet_wrap(~ Election.Year.) 

#is there a correlation between probation and prison

cor.test(data = CA_Merged_Data , 
         ~ Percentage_Prison + Percentage_Probation)

#graph above


CA_Merged_Data %>%
  ggplot(aes(color = Decarceratory. , x = Prison , y = Probation)) + geom_point()

```


Analyzing PA Data
```{r}

PA_Prison_Sentences_by_Decarceratory <- aov(data = PA_Merged_Data , Percentage_Prison ~ Decarceratory)

summary(PA_Prison_Sentences_by_Decarceratory)


TukeyHSD(PA_Prison_Sentences_by_Decarceratory)


wilcox.test(PA_Merged_Data %>%
              filter(Decarceratory == 1) %>%
              select(Percentage_Prison) %>%
              na.omit %>%
              unlist, 
            PA_Merged_Data %>%
              filter(Decarceratory == 0) %>%
              select(Percentage_Prison) %>%
              na.omit %>%
              unlist)

#still comparing medians

med1way(data = PA_Merged_Data ,
        Percentage_Prison ~ Decarceratory)

med2way(data = PA_Merged_Data ,
        Percentage_Prison ~ Decarceratory + Election_Year)

#boxplot

PA_Merged_Data %>%
  ggplot(aes(fill = Election_Year, x = Decarceratory , y = Percentage_Prison)) + geom_boxplot() + facet_wrap(~ Election_Year) 

#is there a correlation between probation and prison

cor.test(data = PA_Merged_Data , 
         ~ Percentage_Prison + Percentage_Probation)

#graph above

PA_Merged_Data %>%
  ggplot(aes(color = Decarceratory , x = State_Prison , y = Probation)) + geom_point()

```
```

VISUALIZATIONS

```{r}


```
